### 2024-09-20 paper/paper.qmd
initial upload
cannot fork twice

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-09-21 paper/paper.qmd
fixed

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Cost of Living in Toronto"
+subtitle: "low-income families can not "
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+  - Hyunje Park
+thanks: "Code and data are available at: https://github.com/davidpxrk/cost-of-living-toronto"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "In summary, all the t."
 format: pdf

### 2024-09-21 scripts/01-download_data.R

@@ -1,26 +1,43 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from OpenDataToronto
+# Author: Hyunje Park
+# Date: 24 September 2024
+# Contact: hyunje.park@mail.utoronto.ca 
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
 
 #### Workspace setup ####
+#install.packages('opendatatoronto')
+#install.packages("dplyr")
+#install.packages('tidyverse')
+#install.packages('sf')
+
 library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
-
+library(dplyr)
+library(sf)
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
 
+# get package
+link <- "52182614-1f0b-42be-aca4-3f86dc8e004c"
 
+package <- show_package(link)
+package
+
+# get all resources for this package
+resources <- list_package_resources(link)
+
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+datastore_resources
+
+# load the first data store resource as a sample
+raw_data <- filter(resources,row_number()==1) %>% get_resource()
+raw_data
 
 #### Save data ####
-# [...UPDATE THIS...]
+
 # change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+write_csv(raw_data, "data/raw_data/raw_data.csv") 
 
          

### 2024-09-21 scripts/02-data_cleaning.R

@@ -1,44 +1,26 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw data of Toronto's cost of living for the parameters
+# Author: Hyunje Park
+# Date: 24 September 2024
+# Contact: hyunje.park@mail.utoronto.ca 
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
 
 #### Workspace setup ####
+#install.packages('opendatatoronto')
+#install.packages("dplyr")
+#install.packages('tidyverse')
+#install.packages('sf')
+
+library(opendatatoronto)
 library(tidyverse)
+library(dplyr)
+library(sf)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+raw_data <- read.csv("data/raw_data/raw_data.csv")
+head(raw_data, 2)
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
 
 #### Save data ####
-write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+#write.csv(cleaned_data, "data/cleaned_data/cleaned_data.csv")

### 2024-09-22 paper/paper.qmd
finished data_cleaning.R

--- 2024-09-20 paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Cost of Living in Toronto"
+subtitle: "Are low-income families able to survive financially here?"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+  - Hyunje Park
+thanks: "Code and data are available at: https://github.com/davidpxrk/cost-of-living-toronto"
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper analyzes the living-cost data from low-income families living in Toronto, which is hosted publicly by the Open Data Toronto.
+It uses information on household income, expenses, government fiscal transfers and benefits to determine if low-income families are able to survive financially in Toronto.
+It was discovered that expenses exceeded income for most families (meaning low-income families are not able to survive in Toronto), and rent, food costs were among the highest expenses.
+This finding is important as it shows cost disprity in expenses, and guide fiscal policymakers on how they can help low-income families with living costs."
 format: pdf
+toc: true
 number-sections: true
@@
 #| message: false
+#install.packages('opendatatoronto')
+#install.packages("dplyr")
+#install.packages('tidyverse')
+#install.packages('sf')
 
+library(palmerpenguins)
+library(opendatatoronto)
 library(tidyverse)
-library(palmerpenguins)
+library(dplyr)
+library(sf)
 ```
@@
 
-# Data {#sec-data}
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
-
-\newpage
-
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
 \newpage

### 2024-09-22 scripts/01-download_data.R

@@ -21,7 +21,7 @@ library(sf)
 # get package
 link <- "52182614-1f0b-42be-aca4-3f86dc8e004c"
 
-package <- show_package(link)
+package <- show_package("52182614-1f0b-42be-aca4-3f86dc8e004c")
 package
 
 # get all resources for this package

### 2024-09-22 scripts/02-data_cleaning.R

@@ -21,6 +21,73 @@ library(sf)
 raw_data <- read.csv("data/raw_data/raw_data.csv")
 head(raw_data, 2)
 
+# remove the last column, it is empty for no reason
+raw_data <- raw_data[1:(length(raw_data)-1)]
 
-#### Save data ####
-#write.csv(cleaned_data, "data/cleaned_data/cleaned_data.csv")
+# replace all the NA values with 0 
+raw_data[is.na(raw_data)] <- 0
+
+# rename Scenarios to Household to fit context better
+colnames(raw_data) = c("Category", "Item", 
+                       "Household_1",
+                       "Household_2",
+                       "Household_3",
+                       "Household_4",
+                       "Household_5",
+                       "Household_6",
+                       "Household_7",
+                       "Household_8",
+                       "Household_9",
+                       "Household_10",
+                       "Household_11")
+
+cleaned_data <- raw_data
+
+#inverted version of raw_data, will need it for graphs in Section 3 Results
+cleaned_data_2 <- data.frame(t(raw_data[-1]))
+colnames(cleaned_data_2) <- raw_data[, 1]
+
+# change column names for inverted dataframe
+colnames(cleaned_data_2) <- c("Income 1",
+                              "Income 2",
+                              "Income 3",
+                              "Income 4",
+                              "Income 5",
+                              "Income 6",
+                              "Income 7",
+                              "Income 8",
+                              "Income 9",
+                              "Expense 1",
+                              "Expense 2",
+                              "Income 10",
+                              "Income 11",
+                              "Total Income",
+                              "Expense 3",
+                              "Expense 4",
+                              "Expense 5",
+                              "Expense 6",
+                              "Total Expenses",
+                              "Remaining Funds",
+                              "Other 2",
+                              "Other 3")
+
+# delete the first row (empty as well, unnecessary)
+cleaned_data_2 <- cleaned_data_2[-1,]
+
+# change all the column datatype back to double
+cleaned_data_2[] <- sapply(cleaned_data_2, as.numeric)
+cleaned_data_2$"Household" <-  c("Household_1",
+                                 "Household_2",
+                                 "Household_3",
+                                 "Household_4",
+                                 "Household_5",
+                                 "Household_6",
+                                 "Household_7",
+                                 "Household_8",
+                                 "Household_9",
+                                 "Household_10",
+                                 "Household_11")
+
[…]

### 2024-09-22 paper/paper.qmd
fixed a launch error in paper.qmd

@@
 
-library(palmerpenguins)
+# RUN THIS IN THE TERMINAL IF THERE'S AN ERROR IN EXPORTING TO PDF
+# Source: https://stackoverflow.com/questions/75906430/why-cant-i-render-any-quarto-documents-into-pdfs
+# quarto install tool tinytex
+
 library(opendatatoronto)
@@
 library(sf)
+
+# Read the 2 cleaned CSV files
+analysis_data = read_csv('data/cleaned_data/cleaned_data.csv')
+analysis_data_inverted = read_csv('data/cleaned_data/cleaned_data_inverted.csv')
+
+analysis_data
+analysis_data_inverted
+
 ```
@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
-
-The remainder of this paper is structured as follows. @sec-data....
-
-\newpage
+testing testing
 

### 2024-09-22 scripts/02-data_cleaning.R

@@ -44,11 +44,11 @@ colnames(raw_data) = c("Category", "Item",
 cleaned_data <- raw_data
 
 #inverted version of raw_data, will need it for graphs in Section 3 Results
-cleaned_data_2 <- data.frame(t(raw_data[-1]))
-colnames(cleaned_data_2) <- raw_data[, 1]
+cleaned_data_inverted <- data.frame(t(raw_data[-1]))
+colnames(cleaned_data_inverted) <- raw_data[, 1]
 
-# change column names for inverted dataframe
-colnames(cleaned_data_2) <- c("Income 1",
+# change all the column names for inverted dataframe
+colnames(cleaned_data_inverted) <- c("Income 1",
                               "Income 2",
                               "Income 3",
                               "Income 4",
@@ -72,11 +72,11 @@ colnames(cleaned_data_2) <- c("Income 1",
                               "Other 3")
 
 # delete the first row (empty as well, unnecessary)
-cleaned_data_2 <- cleaned_data_2[-1,]
+cleaned_data_inverted <- cleaned_data_inverted[-1,]
 
 # change all the column datatype back to double
-cleaned_data_2[] <- sapply(cleaned_data_2, as.numeric)
-cleaned_data_2$"Household" <-  c("Household_1",
+cleaned_data_inverted[] <- sapply(cleaned_data_inverted, as.numeric)
+cleaned_data_inverted$"Household" <-  c("Household_1",
                                  "Household_2",
                                  "Household_3",
                                  "Household_4",
@@ -90,4 +90,5 @@ cleaned_data_2$"Household" <-  c("Household_1",
 
 # Save data
 write_csv(cleaned_data, 'data/cleaned_data/cleaned_data.csv')
-write_csv(cleaned_data_2, 'data/cleaned_data/cleaned_data_2.csv')
+write_csv(cleaned_data_inverted, 'data/cleaned_data/cleaned_data_inverted.csv')
+

### 2024-09-22 paper/paper.qmd
fixed typos in introduction of paper.qmd

@@
 It uses information on household income, expenses, government fiscal transfers and benefits to determine if low-income families are able to survive financially in Toronto.
-It was discovered that expenses exceeded income for most families (meaning low-income families are not able to survive in Toronto), and rent, food costs were among the highest expenses.
-This finding is important as it shows cost disprity in expenses, and guide fiscal policymakers on how they can help low-income families with living costs."
+It was discovered that expenses exceeded income for most families (meaning low-income families were not able to make ends meet in Toronto), and rent, food costs were among the highest expenses.
+This finding is important as it shows cost disparity in expenses, and provide insights to fiscal policymakers on how to better support low-income families with their living costs."
 format: pdf
@@
 
-# Read the 2 cleaned CSV files
-analysis_data = read_csv('data/cleaned_data/cleaned_data.csv')
-analysis_data_inverted = read_csv('data/cleaned_data/cleaned_data_inverted.csv')
-
-analysis_data
-analysis_data_inverted
-
+## Read the 2 cleaned CSV files ##
+analysis_data = read_csv(here::here('data/cleaned_data/cleaned_data.csv'))
+analysis_data_inverted = read_csv(here::here('data/cleaned_data/cleaned_data_inverted.csv'))
 ```
@@
 
+# Data
+
+## Data Source
+
+## Overview of the Dataset
+
+## Average Income of the 11 Households
+
+## Average Expenses of the 11 Households
+
+## Difference between Income and Expense
+
+# Results
+
+# Discussion
+
+# Conclusion
+
+# Appendix
+
+# Statement on LLMs
+
+LLMs and other generative AI tools were not used in the making of this paper.
+
 # References

### 2024-09-22 paper/paper.qmd
fixed issue with knitr package

@@
 #install.packages('sf')
+#remove.packages('knitr')
+install.packages('xfun')
+install.packages('knitr')
 
@@
 
+library(here)
+library(knitr)
 library(opendatatoronto)
@@
 analysis_data = read_csv(here::here('data/cleaned_data/cleaned_data.csv'))
+
 analysis_data_inverted = read_csv(here::here('data/cleaned_data/cleaned_data_inverted.csv'))
@@
 
-testing testing
+Toronto is the biggest and the most populous city in Canada.
+As the fourth largest city in North America, Toronto serves as a catalyst when it comes to being an international hub for many businesses, arts, sports and diverse cultures, attracting people from all around the world to come to Toronto to study at some of the best institutions, work or even simply travel as a Tourist.
+More importantly, Toronto is the most diverse city in Canada, with many people of different cultures, backgrounds, ethnicity coexist, and bridge cultural gaps and create a unity.
+
+While this may seem like a perfect place to live for many international talents, residents of Toronto have said this to be far from the truth, with many sturggling financially to meet ends.
 
 # Data
+```{r}
 
+analysis_data
+#kable(analysis_data)
+#knitr::kable(head(analysis_data[, 1:4]), format = "pipe")
+analysis_data %>% head() %>% knitr::kable()
+```
 ## Data Source

### 2024-09-22 paper/paper.qmd
added introduction

@@
 #remove.packages('knitr')
-install.packages('xfun')
-install.packages('knitr')
+#install.packages('xfun')
+#install.packages('knitr')
 
@@
 ## Read the 2 cleaned CSV files ##
-analysis_data = read_csv(here::here('data/cleaned_data/cleaned_data.csv'))
+cleaned_data = read_csv(here::here('data/cleaned_data/cleaned_data.csv'))
 
-analysis_data_inverted = read_csv(here::here('data/cleaned_data/cleaned_data_inverted.csv'))
+cleaned_data_inverted = read_csv(here::here('data/cleaned_data/cleaned_data_inverted.csv'))
 ```
@@
 
-While this may seem like a perfect place to live for many international talents, residents of Toronto have said this to be far from the truth, with many sturggling financially to meet ends.
+While Toronto may seem like a perfect place to live, residents have said this to be far from the truth, with many sturggling financially to meet ends, especially for those that come from a low income background.
+For example, according to the Daily Bread Food Bank, one in 10 people in Toronto relied on food banks in 2023, a huge increase from one in 20 from 2022 (CITATION HERE).
+
+According to the City of Toronto, the threshold for LIM_AT (threshold to be considered low income) was $26,503 for 1 person household, $37,480 for two person household and $53,005 for a four person household (CITATION, CITY OF TORONTO).
+13.2% of Toronto's population fell under this category of low-income.
+Furthermore, Toronto continues to have a higher rate of low income than any other regions in the GTHA at 8.09% (CITATION HERE), highlighting an issue of income discrepancy.
+
+This paper aims to analyze if low income families living in Toronto are financially able to survive.
+The analysis focuses on two different metrics; income (which includes income from employment, government benefits and transfers such as GST/HST credit, Canada Pension Plan and Canada Worker Benefits) and expense (which includes rent, childcare, food and transportation costs).
+The difference between income and expense can assess whether or not low-income families are able to survive, and can guide future policymakers on what areas of expenses the low-income families can get additional support.
+
+The paper is organized as follows: Section 2 (Data) of this paper introduces the raw dataset of interest and some of the initial observations.
+Section 3 (Results) highlights some of the key insights that were derived from the data analysis.
+Section 4 (Discussion) discusses the observations from Section 3 and draws a conclusion from the findings.
+Finally, Section 5 (Conclusion) draws a brief summary of the key findings.
 
@@
 ```{r}
+#| label: low-income-families
+#| fig-cap: "Table 1: Sample of the Income/Expenses of Low Income Households in Toronto"
+#| echo: false
 
-analysis_data
-#kable(analysis_data)
-#knitr::kable(head(analysis_data[, 1:4]), format = "pipe")
-analysis_data %>% head() %>% knitr::kable()
+# PREVIEW OF THE DATA
+cleaned_data[, 1:5] %>% head() %>% knitr::kable()
 ```
@@
 # References
+
+https://www.toronto.ca/wp-content/uploads/2022/07/9877-City-Planning-2021-Census-Backgrounder-Families-Hhlds-Marital-Status-Income.pdf
+
+https://www.dailybread.ca/wp-content/uploads/2023/11/DB-WhosHungryReport-2023-Digital.pdf

### 2024-09-22 paper/paper.qmd
added all the necessary graphs in paper.qmd

@@
 # Data
+
+This report uses the "About Cost of living in Toronto for low-income households" (CITATION HERE) dataset, which is published publicly by the City of Toronto on the Open Data Portal.
+This data contains information on income such as employment income, GST/HST credits, pension benefits and expenses such as rent, food, transportations for 11 different scenarios of a low-income household.
+The 11 different scenarios of low-income household were manually selected by the City of Toronto to reflect different household characteristics that can increase the reliance on social assistance, such as type of income, single-parent household, etc.
+A detailed explanation of this 11 household scenarios can be found in the Appendix.
+
+This dataset was downloaded and analyzed using the R Programming Language (R Core Team 2022), opendatatoronto (CITATION HERE), knitr (CITATION HERE), tidyverse (CITATION HERE), here (CITATION), dplyr (CITATION) and sf (CITATION HERE).
+
+## Overview of the Dataset
 ```{r}
@@
 ```
-## Data Source
-
-## Overview of the Dataset
 
 ## Average Income of the 11 Households
+```{r}
+#| label: low-income-families
+#| fig-cap: "Table 2: Average Income of Low Income Households in Toronto"
+#| echo: false
+
+avg_income <- filter(cleaned_data, Category == "Income")
+
+# take the average
+avg_income$"Average Income in $" <- rowMeans(subset(avg_income, select = c("Household_1",
+                                                                           "Household_2",
+                                                                           "Household_3",
+                                                                           "Household_4",
+                                                                           "Household_5",
+                                                                           "Household_6",
+                                                                           "Household_7",
+                                                                           "Household_8",
+                                                                           "Household_9",
+                                                                           "Household_10",
+                                                                           "Household_11"
+                                                                           )))
+avg_income <- avg_income %>% select("Item", "Average Income in $")
+
+# rename the columns to something more presentable
+colnames(avg_income) <- c("Type of Income", "Average Income in $")
+
+#round to the nearest integer
+avg_income$"Average Income in $" <- round(avg_income$"Average Income in $", 0)
+
+avg_income %>% head() %>% knitr::kable()
+```
 
 ## Average Expenses of the 11 Households
+```{r}
+#| label: low-income-families
+#| fig-cap: "Table 3: Average Expenses of Low Income Households in Toronto"
+#| echo: false
 
+avg_expense <- filter(cleaned_data, Category == "Employment Deduction" |
+                                    Category == "Selected Expenses")
+
+# take the average of the expenses
+avg_expense$"Average Expense in $" <- rowMeans(subset(avg_expense, select = c("Household_1",
+                                                                           "Household_2",
+                                                                           "Household_3",
+                                                                           "Household_4",
+                                                                           "Household_5",
+                                                                           "Household_6",
+                                                                           "Household_7",
+                                                                           "Household_8",
+                                                                           "Household_9",
+                                                                           "Household_10",
+                                                                           "Household_11"
+                                                                           )))
+avg_expense <- avg_expense %>% select("Item", "Average Expense in $")
+
+# rename the columns to something more presentable
+colnames(avg_expense) <- c("Type of Income", "Average Expense in $")
+
+#round to the nearest integer
+avg_expense$"Average Expense in $" <- round(avg_expense$"Average Expense in $", 0)
+
+# absolute value (because employment insurance is relative to income)
+avg_expense$"Average Expense in $" <- abs(avg_expense$"Average Expense in $") 
+avg_expense %>% head() %>% knitr::kable()
+
+```
 ## Difference between Income and Expense
+```{r}
+#| label: low-income-families
+#| fig-cap: "Table 4: Average Expenses of Low Income Households in Toronto"
+#| echo: false
+
+income_expense_merged <- cleaned_data_inverted
+
+income_expense_merged <- income_expense_merged %>% select("Total Income", "Total Expenses", "Remaining Funds")
+
+income_expense_merged  %>% knitr::kable()
+```
 
 # Results
+```{r}
+#| label: low-income-families
+#| fig-cap: "Table 5: Average Expenses of Low Income Households in Toronto"
+#| echo: false
 
+data <- data.frame(
+  Household = c("Household_1", "Household_2", "Household_3", "Household_4", 
+                "Household_5", "Household_6", "Household_7", "Household_8", 
+                "Household_9", "Household_10", "Household_11"),
+  Income = income_expense_merged$"Total Income",
+  Expenses = income_expense_merged$"Total Expenses"
+)
+
+data_long <- data %>%
+  pivot_longer(cols = c(Income, Expenses), 
+               names_to = "Type", 
+               values_to = "Amount")
+
+ggplot(data_long, aes(x = Household, y = Amount, fill = Type)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Total Income vs Total Expenses for Each Household", 
+       x = "Household", 
+       y = "Amount") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+```{r}
+#| label: low-income-families
+#| fig-cap: "Table 6: Average Expenses of Low Income Households in Toronto"
+#| echo: false
+
+expense_disparity <- head(avg_expense, 6)
+avg_expense
+pie(expense_disparity$'Average Expense in $', expense_disparity$'Type of Income')
+
+```
 # Discussion

### 2024-09-23 paper/paper.qmd
fixed download_data.R to download the correct file

@@
 
-This report uses the "About Cost of living in Toronto for low-income households" (CITATION HERE) dataset, which is published publicly by the City of Toronto on the Open Data Portal.
-This data contains information on income such as employment income, GST/HST credits, pension benefits and expenses such as rent, food, transportations for 11 different scenarios of a low-income household.
+This report uses the "About Cost of living in Toronto for low-income households" (CITATION HERE) dataset, which is published publicly by the Social Development, Finance & Administration of Toronto on City of Toronto's Open Data Portal.
+This data contains information on income such as employment income, GST/HST credits, pension benefits and expenses such as rent, food, transportation costs for 11 different scenarios of a low-income household.
 The 11 different scenarios of low-income household were manually selected by the City of Toronto to reflect different household characteristics that can increase the reliance on social assistance, such as type of income, single-parent household, etc.
 A detailed explanation of this 11 household scenarios can be found in the Appendix.
+The dataset is also within a monthly time period (monthly income and monthly expenses)
 
-This dataset was downloaded and analyzed using the R Programming Language (R Core Team 2022), opendatatoronto (CITATION HERE), knitr (CITATION HERE), tidyverse (CITATION HERE), here (CITATION), dplyr (CITATION) and sf (CITATION HERE).
+The dataset was downloaded and analyzed using the R Programming Language (R Core Team 2022), opendatatoronto (CITATION HERE), knitr (CITATION HERE), tidyverse (CITATION HERE), here (CITATION), dplyr (CITATION) and sf (CITATION HERE).
 
 ## Overview of the Dataset
+
+The dataset published by Social Development, Finance & Administration of Toronto was last updated on August 2024 on an annual basis.
+However, the dataset uses price estimates from July 2022 - July 2023 for its income and expenses of families.
+The variables chosen for this analysis were the 11 households, for each entry of the income and expense types.
+See Table 1 for an overview.
+
 ```{r}
@@
 
+It is also to note that there were differing datasets based on different rent estimates.
+The dataset for rent prices using Toronto Regional Real Estate Board Market Rent was chosen.
+Furthermore, there were several entries in the dataset that were irrelevant, such as percentage of rent on income, which were filtered out.
+
 ## Average Income of the 11 Households
+
+By far, income from employment was the major income source for all the 11 households in Toronto, at approximately $1603 per month.
+This number also falls shy below the "low-income" threshold set by the City of Toronto, which is approximately $2208.5 (CITY OF TORONTO CITATION).
+See Table 2 for a sample of the dataset.
+
 ```{r}
@@
 
+The numbers may not add up as they were rounded to the nearest whole number.
+
 ## Average Expenses of the 11 Households
+
+There is a huge price disparity between the expenses of the 11 families. The biggest expense was 
 ```{r}
@@
 income_expense_merged <- cleaned_data_inverted
+
+cleaned_data_inverted
 

### 2024-09-24 paper/paper.qmd
fixed a typo in data

@@
 The dataset published by Social Development, Finance & Administration of Toronto was last updated on August 2024 on an annual basis.
-However, the dataset uses price estimates from July 2022 - July 2023 for its income and expenses of families.
+However, the dataset uses price estimates from July 2022 - July 2023 for its income and expenses of the families.
 The variables chosen for this analysis were the 11 households, for each entry of the income and expense types.
 See Table 1 for an overview.
+It is noted that expenses and households 4 to 11 are not visible in the preview.
 
@@
 This number also falls shy below the "low-income" threshold set by the City of Toronto, which is approximately $2208.5 (CITY OF TORONTO CITATION).
-See Table 2 for a sample of the dataset.
+See Table 2 for a overview of the dataset.
 
@@
 
-There is a huge price disparity between the expenses of the 11 families. The biggest expense was 
+There is a huge price disparity between the expenses of the 11 families.
+The biggest expense was rent by a huge margin, at an average of $3146 per month, followed by food costs at $705.
+The huge spike in rent highlights a need for further investigation.
+See Table 3 for an overview of the dataset.
+
 ```{r}
@@
 # rename the columns to something more presentable
-colnames(avg_expense) <- c("Type of Income", "Average Expense in $")
+colnames(avg_expense) <- c("Type of Expense", "Average Expense in $")
 
@@
 ## Difference between Income and Expense
+
+To better understand the relationship between the income and expenses of the 11 families, they were merged together into a single dataset, as shown in Table 4.
+In addition to this, a variable for their difference was also created.
+
 ```{r}
@@
 # Results
+
+The distribution of income and expense for each household, shown in Figure 1 shows
 ```{r}
 #| label: low-income-families
-#| fig-cap: "Table 5: Average Expenses of Low Income Households in Toronto"
+#| fig-cap: "Figure 5: Average Expenses of Low Income Households in Toronto"
 #| echo: false
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
 ```
@@
 ```
+
 # Discussion
+
+The tables and graphs above show that the expenses of these families exceed far of the income, suggesting they are not financially able to survive in Toronto.
+However, this raises the question, would providing more income to these families solve their financial issues?
+While giving more money to these families certainly help, having more money does not necessarily mean these families are utilizing their funds in an efficient way.
+For example, rent could differ significantly in price if the household chooses to reside in a highly populated area and contained more rooms.
+Furthermore, food prices can be significantly more if the families chose to eat out, rather than utilizing grocery stores.
+
+Without context to this information, determining the efficiency of spending of the families is impossible.
+However, looking at the results show some inefficiencies in spending.
+For example, in Figure 3, the average rent was $3146 per month, a big difference from the average rent for a one-bedroom condo of $2,529 (Toronto Regional Estate Board, 2024).
+Expenses such as this could suggest that expenses could be correlated with poor financial decisions, however more research must be conducted to draw those conclusions.
 
@@
 # Appendix
+
+## A.1 Detailed Information on the 11 Households
+
+The 11 different scenarios of low-income household were manually selected by the City of Toronto to reflect different household characteristics that can increase the reliance on social assistance, such as type of income, single-parent household, etc.
+The structure of the 11 households is as follows:
+
+Family 1:
 

### 2024-09-24 paper/paper.qmd
fixed typos in paper.qmd

@@
 abstract: "This paper analyzes the living-cost data from low-income families living in Toronto, which is hosted publicly by the Open Data Toronto.
-It uses information on household income, expenses, government fiscal transfers and benefits to determine if low-income families are able to survive financially in Toronto.
+It uses information on household income, expenses, government fiscal transfers, grants and benefits to determine if low-income families are able to survive financially in Toronto.
 It was discovered that expenses exceeded income for most families (meaning low-income families were not able to make ends meet in Toronto), and rent, food costs were among the highest expenses.
-This finding is important as it shows cost disparity in expenses, and provide insights to fiscal policymakers on how to better support low-income families with their living costs."
+This finding is important as it shows cost disparity in expenses, and provide insights to fiscal policymakers on what areas to better support low-income families with their living costs."
 format: pdf
@@
 Toronto is the biggest and the most populous city in Canada.
-As the fourth largest city in North America, Toronto serves as a catalyst when it comes to being an international hub for many businesses, arts, sports and diverse cultures, attracting people from all around the world to come to Toronto to study at some of the best institutions, work or even simply travel as a Tourist.
+As the fourth largest city in North America, Toronto serves as a catalyst when it comes to being an international hub for many businesses, arts, sports and diverse cultures, attracting people from all around the world to come to Toronto to study at some of the best institutions, work or even simply travel as a tourist.
 More importantly, Toronto is the most diverse city in Canada, with many people of different cultures, backgrounds, ethnicity coexist, and bridge cultural gaps and create a unity.
 
-While Toronto may seem like a perfect place to live, residents have said this to be far from the truth, with many sturggling financially to meet ends, especially for those that come from a low income background.
-For example, according to the Daily Bread Food Bank, one in 10 people in Toronto relied on food banks in 2023, a huge increase from one in 20 from 2022 (CITATION HERE).
+While Toronto may seem like a perfect place to live, residents have said this to be far from the truth, with many struggling financially to meet ends, especially for those from a low income background.
+For example, according to the Daily Bread Food Bank, 1/10 people in Toronto relied on food banks in 2023, a huge increase from 1/20 from 2022 (CITATION HERE).
 
@@
 This paper aims to analyze if low income families living in Toronto are financially able to survive.
-The analysis focuses on two different metrics; income (which includes income from employment, government benefits and transfers such as GST/HST credit, Canada Pension Plan and Canada Worker Benefits) and expense (which includes rent, childcare, food and transportation costs).
-The difference between income and expense can assess whether or not low-income families are able to survive, and can guide future policymakers on what areas of expenses the low-income families can get additional support.
+The analysis focuses on two different metrics; income (which includes income from employment, government benefits and transfers such as GST/HST credit, Canada Pension Plan and Canada Worker Benefits) and expenses (which includes rent, childcare, food and transportation costs).
+The difference between income and expenses can assess whether or not low-income families are able to survive, and can guide future policymakers on what areas of expenses the low-income families can get additional support.
 
@@
 A detailed explanation of this 11 household scenarios can be found in the Appendix.
-The dataset is also within a monthly time period (monthly income and monthly expenses)
+The dataset is also within a monthly time period, reflecting monthly income and expenses.
 
@@
 However, the dataset uses price estimates from July 2022 - July 2023 for its income and expenses of the families.
-The variables chosen for this analysis were the 11 households, for each entry of the income and expense types.
+The variables chosen for this analysis were variables for the 11 households, for each entry of the income and expense types.
 See Table 1 for an overview.
-It is noted that expenses and households 4 to 11 are not visible in the preview.
+It is noted that expenses and Households 4 to 11 are not visible in the preview, and expenses are not visible due to the long list of income that is above it.
 
@@
 
-It is also to note that there were differing datasets based on different rent estimates.
-The dataset for rent prices using Toronto Regional Real Estate Board Market Rent was chosen.
+It is also to note that the dataset for rent prices using Toronto Regional Real Estate Board Market Rent was chosen.
 Furthermore, there were several entries in the dataset that were irrelevant, such as percentage of rent on income, which were filtered out.
@@
 
-By far, income from employment was the major income source for all the 11 households in Toronto, at approximately $1603 per month.
+Table 2 shows that income from employment was the major income source for all the 11 households in Toronto by a large margin, at approximately $1603 per month.
 This number also falls shy below the "low-income" threshold set by the City of Toronto, which is approximately $2208.5 (CITY OF TORONTO CITATION).
-See Table 2 for a overview of the dataset.
 
 ```{r}
-#| label: low-income-families
+#| label: low-income-families2
 #| fig-cap: "Table 2: Average Income of Low Income Households in Toronto"
@@
 
-The numbers may not add up as they were rounded to the nearest whole number.
+The numbers may not add up properly as they were rounded to the nearest whole number.
 
@@
 There is a huge price disparity between the expenses of the 11 families.
-The biggest expense was rent by a huge margin, at an average of $3146 per month, followed by food costs at $705.
-The huge spike in rent highlights a need for further investigation.
+The biggest expense was rent, at an average of $3146 per month, followed by food costs at $705.
+The huge cost rent highlights a need for further investigation and is discussed in Section 4.
 See Table 3 for an overview of the dataset.
@@
 ```{r}
-#| label: low-income-families
+#| label: low-income-families3
 #| fig-cap: "Table 3: Average Expenses of Low Income Households in Toronto"
@@
 
-To better understand the relationship between the income and expenses of the 11 families, they were merged together into a single dataset, as shown in Table 4.
+To better understand the relationship between the income and expenses of the 11 families, they were merged together into a single dataset.
 In addition to this, a variable for their difference was also created.
+See Table 4 for a preview of the dataset.
 
 ```{r}
-#| label: low-income-families
+#| label: low-income-families4
 #| fig-cap: "Table 4: Average Expenses of Low Income Households in Toronto"
@@
 income_expense_merged <- cleaned_data_inverted
-
-cleaned_data_inverted
 
@@
 
-The distribution of income and expense for each household, shown in Figure 1 shows
+Figure 1 shows income and expense for each household. Overall, expenses exceed income for all households. Only one household had an income that exceed expenses.
+
 ```{r}
-#| label: low-income-families
+#| label: low-income-families5
 #| fig-cap: "Figure 5: Average Expenses of Low Income Households in Toronto"
@@
 ```
+Furthermore, Figure 2 gives us a better understanding of the spending behind the high expenses. Rent was by far the most expensive, comprising X% of the expenses.
+
 ```{r}
-#| label: low-income-families
-#| fig-cap: "Table 6: Average Expenses of Low Income Households in Toronto"
+#| label: low-income-families6
+#| fig-cap: "Figure 2: Average Expenses of Low Income Households in Toronto"
 #| echo: false
@@
 expense_disparity <- head(avg_expense, 6)
-avg_expense
 pie(expense_disparity$'Average Expense in $', expense_disparity$'Type of Income')
 
+percentages <- round(100 * expense_disparity$'Average Expense in $' / sum(expense_disparity$'Average Expense in $'), 1)
+
+# Create labels with percentages
+labels <- paste(expense_disparity$'Type of Income', percentages, "%")
+
+# Generate the pie chart
+pie(expense_disparity$'Average Expense in $')
+text(x = 0, y = 0, labels = labels, pos = 3, cex = 0.8, col = "white")
 ```
@@
 The tables and graphs above show that the expenses of these families exceed far of the income, suggesting they are not financially able to survive in Toronto.
+Furthermore, rent comprised X% of the total expenses, highlighting a need for policymakers to focus on housing prices.
 However, this raises the question, would providing more income to these families solve their financial issues?
-While giving more money to these families certainly help, having more money does not necessarily mean these families are utilizing their funds in an efficient way.
-For example, rent could differ significantly in price if the household chooses to reside in a highly populated area and contained more rooms.
-Furthermore, food prices can be significantly more if the families chose to eat out, rather than utilizing grocery stores.
+While giving more money to these families certainly help, but having more money does not necessarily mean these families are utilizing their funds in an efficient way.
+For example, rent prices could differ significantly if a family chooses to reside in a highly populated area and if the house contained more rooms.
+Furthermore, food prices can be significantly more if the families chose to eat out, rather than utilizing lower-costing grocery stores.
 
 Without context to this information, determining the efficiency of spending of the families is impossible.
-However, looking at the results show some inefficiencies in spending.
-For example, in Figure 3, the average rent was $3146 per month, a big difference from the average rent for a one-bedroom condo of $2,529 (Toronto Regional Estate Board, 2024).
+However, looking at the results suggest some inefficiencies in spending.
+For example, in Figure 3, the average rent was $3146 per month for the 11 families, a big difference from the average rent for a one-bedroom condo of $2,529 in Toronto for Q2 of 2023 (Toronto Regional Estate Board, 2024).
 Expenses such as this could suggest that expenses could be correlated with poor financial decisions, however more research must be conducted to draw those conclusions.
@@
 # Conclusion
+
+In summary, expenses exceeded income for most of the 11 household scenarios generated by the City of Toronto, demonstrating low-income families may not be able to make ends meet financially.
+Furthermore, rent made up a significant portion of the families' expenses, guiding policymakers to focus on this area.
+Further research on consumer spending behavior could better explain the expenses behind the families.
 

### 2024-09-24 paper/paper.qmd
added cross references to graphs

@@
 While Toronto may seem like a perfect place to live, residents have said this to be far from the truth, with many struggling financially to meet ends, especially for those from a low income background.
-For example, according to the Daily Bread Food Bank, 1/10 people in Toronto relied on food banks in 2023, a huge increase from 1/20 from 2022 (CITATION HERE).
+For example, according to the Daily Bread Food Bank, `1/10` people in Toronto relied on food banks in 2023, a huge increase from `1/20` from 2022 (CITATION HERE).
 
-According to the City of Toronto, the threshold for LIM_AT (threshold to be considered low income) was $26,503 for 1 person household, $37,480 for two person household and $53,005 for a four person household (CITATION, CITY OF TORONTO).
-13.2% of Toronto's population fell under this category of low-income.
-Furthermore, Toronto continues to have a higher rate of low income than any other regions in the GTHA at 8.09% (CITATION HERE), highlighting an issue of income discrepancy.
+According to the City of Toronto, the threshold for `LIM_AT` (threshold to be considered low income) was `$26,503` for 1 person household, `$37,480` for two person household and `$53,005` for a four person household (CITATION, CITY OF TORONTO).
+`13.2%` of Toronto's population fell under this category of low-income.
+Furthermore, Toronto continues to have a higher rate of low income than any other regions in the GTHA at `8.09%` (CITATION HERE), highlighting an issue of income discrepancy.
 
@@
 
-The paper is organized as follows: Section 2 (Data) of this paper introduces the raw dataset of interest and some of the initial observations.
-Section 3 (Results) highlights some of the key insights that were derived from the data analysis.
-Section 4 (Discussion) discusses the observations from Section 3 and draws a conclusion from the findings.
-Finally, Section 5 (Conclusion) draws a brief summary of the key findings.
+The paper is organized as follows: @sec-data (Data) of this paper introduces the raw dataset of interest and some of the initial observations.
+@sec-results (Results) highlights some of the key insights that were derived from the data analysis.
+@sec-discussion (Discussion) discusses the observations from Section 3 and draws a conclusion from the findings.
+Finally, @sec-conclusion (Conclusion) draws a brief summary of the key findings.
 
-# Data
+# Data {#sec-data}
 
@@
 
-The dataset published by Social Development, Finance & Administration of Toronto was last updated on August 2024 on an annual basis.
+The dataset published by `Social Development, Finance & Administration of Toronto` was last updated on August 2024 on an annual basis.
 However, the dataset uses price estimates from July 2022 - July 2023 for its income and expenses of the families.
@@
 
-It is also to note that the dataset for rent prices using Toronto Regional Real Estate Board Market Rent was chosen.
-Furthermore, there were several entries in the dataset that were irrelevant, such as percentage of rent on income, which were filtered out.
+It is also to note that the dataset for rent prices using `Toronto Regional Real Estate Board Market Rent` was chosen.
+Furthermore, there were several entries in the dataset that were irrelevant, such as `percentage of rent on income`, which were filtered out.
 
@@
 
-Table 2 shows that income from employment was the major income source for all the 11 households in Toronto by a large margin, at approximately $1603 per month.
+`Table 2` shows that income from employment was the major income source for all the 11 households in Toronto by a large margin, at approximately $1603 per month.
 This number also falls shy below the "low-income" threshold set by the City of Toronto, which is approximately $2208.5 (CITY OF TORONTO CITATION).
@@
 The biggest expense was rent, at an average of $3146 per month, followed by food costs at $705.
-The huge cost rent highlights a need for further investigation and is discussed in Section 4.
-See Table 3 for an overview of the dataset.
+The huge cost rent highlights a need for further investigation and is discussed in @sec-discussion.
+See `Table 3` for an overview of the dataset.
 
@@
 In addition to this, a variable for their difference was also created.
-See Table 4 for a preview of the dataset.
+See `Table 4` for a preview of the dataset.
 
@@
 
-# Results
+# Results {#sec-results}
 
-Figure 1 shows income and expense for each household. Overall, expenses exceed income for all households. Only one household had an income that exceed expenses.
+`Figure 1` shows income and expense for each household. Overall, expenses exceed income for all households. Only one household had an income that exceed expenses.
 
@@
 ```
-Furthermore, Figure 2 gives us a better understanding of the spending behind the high expenses. Rent was by far the most expensive, comprising X% of the expenses.
+Furthermore, `Figure 2` gives us a better understanding of the spending behind the high expenses. Rent was by far the most expensive, comprising X% of the expenses.
 
@@
 #| label: low-income-families6
-#| fig-cap: "Figure 2: Average Expenses of Low Income Households in Toronto"
+#| fig-cap: "Figure 2: Expense Breakdown of Low Income Households in Toronto"
 #| echo: false
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 However, looking at the results suggest some inefficiencies in spending.
-For example, in Figure 3, the average rent was $3146 per month for the 11 families, a big difference from the average rent for a one-bedroom condo of $2,529 in Toronto for Q2 of 2023 (Toronto Regional Estate Board, 2024).
+For example, in `Figure 3`, the average rent was $3146 per month for the 11 families, a big difference from the average rent for a one-bedroom condo of $2,529 in Toronto for Q2 of 2023 (Toronto Regional Estate Board, 2024).
 Expenses such as this could suggest that expenses could be correlated with poor financial decisions, however more research must be conducted to draw those conclusions.
 
-# Conclusion
+# Conclusion {#sec-conclusion}
 
-In summary, expenses exceeded income for most of the 11 household scenarios generated by the City of Toronto, demonstrating low-income families may not be able to make ends meet financially.
+In summary, expenses exceeded income for most of the 11 household scenarios generated by the `City of Toronto`, demonstrating low-income families may not be able to make ends meet financially.
 Furthermore, rent made up a significant portion of the families' expenses, guiding policymakers to focus on this area.
@@
 
-# Appendix
+# Appendix {#sec-appendix}
 
@@
 
-The 11 different scenarios of low-income household were manually selected by the City of Toronto to reflect different household characteristics that can increase the reliance on social assistance, such as type of income, single-parent household, etc.
+The 11 different scenarios of low-income household were manually selected by the `City of Toronto` to reflect different household characteristics that can increase the reliance on social assistance, such as type of income, single-parent household, etc.
 The structure of the 11 households is as follows:
 
-Family 1:
+Household 1: 2 adults (male and female ages 31-50) with no income, 2 children (unable to contribute to income).
+Household 2: 2 adults (male and female ages 31-50) with minimum wage pay ($15.00/hr), 2 children (unable to contribute to income).
+Household 3: 2 adults (male and female ages 31-50) with median income, 2 children (unable to contribute to income).
+Household 4: 1 adult (female age 31-50) with no income, 2 children (unable to contribute to income).
+Household 5: 1 adult (male age 31-50) with no income.
+Household 6: 1 adult (male age 31-50) under `Ontario Disability Support Program`, 2 children (unable to contribute to income).
+Household 7: 1 adult (female age 70+) under Old Age Security/Guaranteed Annual Income Supplement.
+Household 8: 2 adults (male and female age 31-50) under Ontario Disability Support Program.
+Household 9: 1 adult (female age 19-30) under Ontario Disability Support Program.
+Household 10: 1 adult (female age 31-50) with minimum wage pay ($15.00/hr) with 2 children (unable to contribute to income).
+Household 11: 1 adult (female age 31-50) with minimum wage pay ($15.00/hr).
+
+The children mentioned are under the mandatory working age limit and therefore unable to join the workforce.
 

### 2024-09-24 paper/paper.qmd
added citations in references.bib

@@
 #install.packages('sf')
-#remove.packages('knitr')
-#install.packages('xfun')
+#remove.packages('knitr')S
 #install.packages('knitr')
+#install.packages('ggplot2)
 
@@
 # quarto install tool tinytex
-
+library(ggplot2)
 library(here)
@@
 While Toronto may seem like a perfect place to live, residents have said this to be far from the truth, with many struggling financially to meet ends, especially for those from a low income background.
-For example, according to the Daily Bread Food Bank, `1/10` people in Toronto relied on food banks in 2023, a huge increase from `1/20` from 2022 (CITATION HERE).
+For example, according to the Daily Bread Food Bank, `1/10` people in Toronto relied on food banks in 2023, a huge increase from `1/20` from 2022 [@citewhoshungry].
 
-According to the City of Toronto, the threshold for `LIM_AT` (threshold to be considered low income) was `$26,503` for 1 person household, `$37,480` for two person household and `$53,005` for a four person household (CITATION, CITY OF TORONTO).
+According to the City of Toronto, the threshold for `LIM_AT` (threshold to be considered low income) was `$26,503` for 1 person household, `$37,480` for two person household and `$53,005` for a four person household [@citecityoftoronto].
 `13.2%` of Toronto's population fell under this category of low-income.
-Furthermore, Toronto continues to have a higher rate of low income than any other regions in the GTHA at `8.09%` (CITATION HERE), highlighting an issue of income discrepancy.
+Furthermore, Toronto continues to have a higher rate of low income than any other regions in the GTHA at `8.09%` [@citecityoftoronto], highlighting an issue of income discrepancy.
 
@@
 
-This report uses the "About Cost of living in Toronto for low-income households" (CITATION HERE) dataset, which is published publicly by the Social Development, Finance & Administration of Toronto on City of Toronto's Open Data Portal.
+This report uses the "About Cost of living in Toronto for low-income households" [@citecityoftoronto] dataset, which is published publicly by the Social Development, Finance & Administration of Toronto on City of Toronto's Open Data Portal.
 This data contains information on income such as employment income, GST/HST credits, pension benefits and expenses such as rent, food, transportation costs for 11 different scenarios of a low-income household.
@@
 
-The dataset was downloaded and analyzed using the R Programming Language (R Core Team 2022), opendatatoronto (CITATION HERE), knitr (CITATION HERE), tidyverse (CITATION HERE), here (CITATION), dplyr (CITATION) and sf (CITATION HERE).
+The dataset was downloaded and analyzed using the `R Programming Language` [@citeR], `opendatatoronto` [@citeopendatatoronto], `knitr` [@citeknitr], `tidyverse` [@citetidyverse], `here` [@citehere], `dplyr` [@citedplyr] and `sf` [@citesf].
 
@@
 See Table 1 for an overview.
-It is noted that expenses and Households 4 to 11 are not visible in the preview, and expenses are not visible due to the long list of income that is above it.
+It is noted that expenses are not visible in `Table 1` due to the long list of income that occupies the first rows.
 
@@
 `Table 2` shows that income from employment was the major income source for all the 11 households in Toronto by a large margin, at approximately $1603 per month.
-This number also falls shy below the "low-income" threshold set by the City of Toronto, which is approximately $2208.5 (CITY OF TORONTO CITATION).
+This number also falls shy below the "low-income" threshold set by the City of Toronto, which is approximately $2208.5 [@citedcityoftoronto].
 
@@
 However, looking at the results suggest some inefficiencies in spending.
-For example, in `Figure 3`, the average rent was $3146 per month for the 11 families, a big difference from the average rent for a one-bedroom condo of $2,529 in Toronto for Q2 of 2023 (Toronto Regional Estate Board, 2024).
+For example, in `Figure 3`, the average rent was $3146 per month for the 11 families, a big difference from the average rent for a one-bedroom condo of $2,529 in Toronto for Q2 of 2023 [@citerentalmarket].
 Expenses such as this could suggest that expenses could be correlated with poor financial decisions, however more research must be conducted to draw those conclusions.
@@
 # Conclusion {#sec-conclusion}
-
-In summary, expenses exceeded income for most of the 11 household scenarios generated by the `City of Toronto`, demonstrating low-income families may not be able to make ends meet financially.
-Furthermore, rent made up a significant portion of the families' expenses, guiding policymakers to focus on this area.
-Further research on consumer spending behavior could better explain the expenses behind the families.
 
@@
 
-Household 1: 2 adults (male and female ages 31-50) with no income, 2 children (unable to contribute to income).
-Household 2: 2 adults (male and female ages 31-50) with minimum wage pay ($15.00/hr), 2 children (unable to contribute to income).
-Household 3: 2 adults (male and female ages 31-50) with median income, 2 children (unable to contribute to income).
-Household 4: 1 adult (female age 31-50) with no income, 2 children (unable to contribute to income).
-Household 5: 1 adult (male age 31-50) with no income.
-Household 6: 1 adult (male age 31-50) under `Ontario Disability Support Program`, 2 children (unable to contribute to income).
-Household 7: 1 adult (female age 70+) under Old Age Security/Guaranteed Annual Income Supplement.
-Household 8: 2 adults (male and female age 31-50) under Ontario Disability Support Program.
-Household 9: 1 adult (female age 19-30) under Ontario Disability Support Program.
-Household 10: 1 adult (female age 31-50) with minimum wage pay ($15.00/hr) with 2 children (unable to contribute to income).
-Household 11: 1 adult (female age 31-50) with minimum wage pay ($15.00/hr).
+`Household 1`: 2 adults (male and female ages 31-50) with no income, 2 children (unable to contribute to income).
+
+`Household 2`: 2 adults (male and female ages 31-50) with minimum wage pay ($15.00/hr), 2 children (unable to contribute to income).
+
+`Household 3`: 2 adults (male and female ages 31-50) with median income, 2 children (unable to contribute to income).
+
+`Household 4`: 1 adult (female age 31-50) with no income, 2 children (unable to contribute to income).
+
+`Household 5`: 1 adult (male age 31-50) with no income.
+
+`Household 6`: 1 adult (male age 31-50) under `Ontario Disability Support Program`, 2 children (unable to contribute to income).
+
+`Household 7`: 1 adult (female age 70+) under Old Age Security/Guaranteed Annual Income Supplement.
+
+`Household 8`: 2 adults (male and female age 31-50) under Ontario Disability Support Program.
+
+`Household 9`: 1 adult (female age 19-30) under Ontario Disability Support Program.
+
+`Household 10`: 1 adult (female age 31-50) with minimum wage pay ($15.00/hr) with 2 children (unable to contribute to income).
+
+`Household 11`: 1 adult (female age 31-50) with minimum wage pay ($15.00/hr).
 
@@
 # References
-
-https://www.toronto.ca/wp-content/uploads/2022/07/9877-City-Planning-2021-Census-Backgrounder-Families-Hhlds-Marital-Status-Income.pdf
-
-https://www.dailybread.ca/wp-content/uploads/2023/11/DB-WhosHungryReport-2023-Digital.pdf

### 2024-09-24 scripts/01-download_data.R

@@ -32,7 +32,7 @@ datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'
 datastore_resources
 
 # load the first data store resource as a sample
-raw_data <- filter(resources,row_number()==1) %>% get_resource()
+raw_data <- filter(resources,row_number()==5) %>% get_resource()
 raw_data
 
 #### Save data ####

### 2024-09-26 paper/paper.qmd
fixed issue with the pie graph

@@
 This report uses the "About Cost of living in Toronto for low-income households" [@citecityoftoronto] dataset, which is published publicly by the Social Development, Finance & Administration of Toronto on City of Toronto's Open Data Portal.
-This data contains information on income such as employment income, GST/HST credits, pension benefits and expenses such as rent, food, transportation costs for 11 different scenarios of a low-income household.
+This data contains 17 variables on income and expenses such as `employment income`, `GST/HST credits`, `pension benefits`, `rent`, `food`, `transportation costs` for 11 different scenarios of a low-income household.
 The 11 different scenarios of low-income household were manually selected by the City of Toronto to reflect different household characteristics that can increase the reliance on social assistance, such as type of income, single-parent household, etc.
@@
 ```{r}
-#| label: low-income-families
+#| label: low-income-families-preview
 #| fig-cap: "Table 1: Sample of the Income/Expenses of Low Income Households in Toronto"
@@
 `Table 2` shows that income from employment was the major income source for all the 11 households in Toronto by a large margin, at approximately $1603 per month.
-This number also falls shy below the "low-income" threshold set by the City of Toronto, which is approximately $2208.5 [@citedcityoftoronto].
+This number also falls shy below the "low-income" threshold set by the City of Toronto, which is approximately $2208.5 [@citecityoftoronto].
 
 ```{r}
-#| label: low-income-families2
+#| label: low-income-families-average-income
 #| fig-cap: "Table 2: Average Income of Low Income Households in Toronto"
@@
 ```{r}
-#| label: low-income-families3
+#| label: low-income-families-average-expense
 #| fig-cap: "Table 3: Average Expenses of Low Income Households in Toronto"
@@
 avg_expense %>% head() %>% knitr::kable()
-
 ```
@@
 income_expense_merged <- cleaned_data_inverted
+income_expense_merged <- income_expense_merged %>% select("Total Income", "Total Expenses", "Remaining Funds")
 
-income_expense_merged <- income_expense_merged %>% select("Total Income", "Total Expenses", "Remaining Funds")
+# bring back household names, deleted for some reason
+income_expense_merged$"Household" <- c("Household_1",
+                                       "Household_2",
+                                       "Household_3",
+                                       "Household_4",
+                                       "Household_5",
+                                       "Household_6",
+                                       "Household_7",
+                                       "Household_8",
+                                       "Household_9",
+                                       "Household_10",
+                                       "Household_11"
+                                       )
+
+# put the household names at the start
+income_expense_merged <- income_expense_merged %>%
+  select("Household", everything())
 
@@
 #| label: low-income-families5
-#| fig-cap: "Figure 5: Average Expenses of Low Income Households in Toronto"
+#| fig-cap: "Average Expenses of Low Income Households in Toronto"
 #| echo: false
@@
   geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Total Income vs Total Expenses for Each Household", 
+  labs(title = "Figure 1: Total Income vs Total Expenses for Each Household", 
        x = "Household", 
@@
 ```
-Furthermore, `Figure 2` gives us a better understanding of the spending behind the high expenses. Rent was by far the most expensive, comprising X% of the expenses.
+Furthermore, `Figure 2` gives us a better understanding of the spending behind the high expenses. Rent was by far the most expensive, comprising 66% of the expenses.
 
@@
 #| label: low-income-families6
-#| fig-cap: "Figure 2: Expense Breakdown of Low Income Households in Toronto"
+#| fig-cap: "Expense Breakdown of Low Income Households in Toronto"
 #| echo: false
@@
 expense_disparity <- head(avg_expense, 6)
-pie(expense_disparity$'Average Expense in $', expense_disparity$'Type of Income')
-
-percentages <- round(100 * expense_disparity$'Average Expense in $' / sum(expense_disparity$'Average Expense in $'), 1)
-
-# Create labels with percentages
-labels <- paste(expense_disparity$'Type of Income', percentages, "%")
-
-# Generate the pie chart
-pie(expense_disparity$'Average Expense in $')
-text(x = 0, y = 0, labels = labels, pos = 3, cex = 0.8, col = "white")
+pie(expense_disparity$'Average Expense in $', expense_disparity$'Type of Expense', main = "Figure 2: Cost Breakdown of Expenses", radius = 1, cex = 0.8) 
 ```
@@
 The tables and graphs above show that the expenses of these families exceed far of the income, suggesting they are not financially able to survive in Toronto.
-Furthermore, rent comprised X% of the total expenses, highlighting a need for policymakers to focus on housing prices.
+Furthermore, rent comprised 66% of the total expenses, highlighting a need for policymakers to focus on housing prices.
 However, this raises the question, would providing more income to these families solve their financial issues?
 While giving more money to these families certainly help, but having more money does not necessarily mean these families are utilizing their funds in an efficient way.
-For example, rent prices could differ significantly if a family chooses to reside in a highly populated area and if the house contained more rooms.
-Furthermore, food prices can be significantly more if the families chose to eat out, rather than utilizing lower-costing grocery stores.
+For example, rent prices could differ significantly if a family chooses to reside in a less populated area and if the house contained less rooms.
+Furthermore, food prices can be significantly less if the families utilized low-costing grocery stores over eating out at restaurants.
 

### 2024-09-26 paper/paper.qmd
added conclusion in paper.qmd

@@
 
+In summary, this paper investigated the income and and expenses of 11 different low-income household scenarios in Toronto.
+It showed that the average expenses far exceeded the average incomes, proving it impossible for low-income families to survive in Toronto.
+Furthermore, rent was the biggest expense, comprising 66% of the total average income, showing that aspect the biggest need for financial aid.
+Further research may be required to investigate into consumer spending behaviour that could help understand the reasoning behind the high expenses.
+
 # Appendix {#sec-appendix}

### 2024-09-26 scripts/01-download_data.R
added test_data.R

@@ -1,10 +1,10 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from OpenDataToronto
+# Purpose: Downloads and saves the data from Open Data Toronto
 # Author: Hyunje Park
 # Date: 24 September 2024
 # Contact: hyunje.park@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
+# Pre-requisites: Packages listed below
 
 #### Workspace setup ####
 #install.packages('opendatatoronto')

### 2024-09-26 scripts/02-data_cleaning.R

@@ -4,7 +4,7 @@
 # Date: 24 September 2024
 # Contact: hyunje.park@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
+# Pre-requisites: 01-download_data.R, packages listed below.
 
 #### Workspace setup ####
 #install.packages('opendatatoronto')

### 2024-09-26 scripts/03-test_data.R

@@ -1,15 +1,66 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw data of Toronto's cost of living for the parameters
+# Author: Hyunje Park
+# Date: 24 September 2024
+# Contact: hyunje.park@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 01-download_data.R, packages listed below.
 
 
 #### Workspace setup ####
-library(tidyverse)
-# [...UPDATE THIS...]
+#install.packages('opendatatoronto')
+#install.packages("dplyr")
+#install.packages('tidyverse')
+#install.packages('sf')
+#install.packages('devtools')
 
-#### Test data ####
+library(tidyverse)
+library(dplyr)
+library(testthat)
+
+# get the cleaned dataset
+cleaned_data = read_csv(here::here('data/cleaned_data/cleaned_data.csv'))
+# cleaned_data_inverted is just an inverted version of cleaned_data, no need to test.
+
+#### Test cleaned_data ####
+
+# test 1 - test that it has all the columns
+test_that("Has all the columns",
+          {expected = c("Category",
+                       "Item",
+                       "Household_1",
+                       "Household_2",
+                       "Household_3",
+                       "Household_4",
+                       "Household_5",
+                       "Household_6",
+                       "Household_7",
+                       "Household_8",
+                       "Household_9",
+                       "Household_10",
+                       "Household_11")
+          expect_equal(colnames(cleaned_data), expected)})
+
+# test that the dataframe contains no NA values
+test_that("No NA values", 
+          {expect_equal(sum(is.na(cleaned_data)), 0)})
+
+# test that all the Income/Expense variables are not negative
+# (because you cant have negative income and expenses)
+test_that("Income and Expense variables are not negative",
+          {
+          expect_true(all(cleaned_data$"Household_1" >= 0))
+          expect_true(all(cleaned_data$"Household_2" >= 0))
+          expect_true(all(cleaned_data$"Household_3" >= 0))
[…]

### 2024-09-26 paper/paper.qmd
added sketches

@@
 #| label: low-income-families4
-#| fig-cap: "Table 4: Average Expenses of Low Income Households in Toronto"
+#| fig-cap: "Table 4: Differenece of Income and Expenses of Low Income Households in Toronto"
 #| echo: false
@@
 #| label: low-income-families5
-#| fig-cap: "Average Expenses of Low Income Households in Toronto"
+#| fig-cap: "Total Income and Expenses of Low Income Households in Toronto"
 #| echo: false
@@
        x = "Household", 
-       y = "Amount") +
+       y = "Amount in $") +
   theme_minimal() +

### 2024-09-27 scripts/03-test_data.R
finished test_data.R

@@ -3,9 +3,7 @@
 # Author: Hyunje Park
 # Date: 24 September 2024
 # Contact: hyunje.park@mail.utoronto.ca 
-# License: MIT
 # Pre-requisites: 01-download_data.R, packages listed below.
-
 
 #### Workspace setup ####
 #install.packages('opendatatoronto')
@@ -20,9 +18,10 @@ library(testthat)
 
 # get the cleaned dataset
 cleaned_data = read_csv(here::here('data/cleaned_data/cleaned_data.csv'))
-# cleaned_data_inverted is just an inverted version of cleaned_data, no need to test.
+cleaned_data_inverted = read_csv(here::here('data/cleaned_data/cleaned_data_inverted.csv'))
 
 #### Test cleaned_data ####
+### NEED TO TEST INCOME/EXPENSE IS PROPER ###
 
 # test 1 - test that it has all the columns
 test_that("Has all the columns",
@@ -45,22 +44,31 @@ test_that("Has all the columns",
 test_that("No NA values", 
           {expect_equal(sum(is.na(cleaned_data)), 0)})
 
-# test that all the Income/Expense variables are not negative
-# (because you cant have negative income and expenses)
-test_that("Income and Expense variables are not negative",
+# test that all the Income sources are positive values
+# (cant have negative Income)
+test_that("All Income variables are not negative",
           {
-          expect_true(all(cleaned_data$"Household_1" >= 0))
-          expect_true(all(cleaned_data$"Household_2" >= 0))
-          expect_true(all(cleaned_data$"Household_3" >= 0))
-          expect_true(all(cleaned_data$"Household_4" >= 0))
-          expect_true(all(cleaned_data$"Household_5" >= 0))
-          expect_true(all(cleaned_data$"Household_6" >= 0))
-          expect_true(all(cleaned_data$"Household_7" >= 0))
-          expect_true(all(cleaned_data$"Household_8" >= 0))
-          expect_true(all(cleaned_data$"Household_9" >= 0))
-          expect_true(all(cleaned_data$"Household_10" >= 0))
-          expect_true(all(cleaned_data$"Household_11" >= 0))
+          expect_true(all(cleaned_data_inverted$"Income 1" >= 0))
+          expect_true(all(cleaned_data_inverted$"Income 2" >= 0))
+          expect_true(all(cleaned_data_inverted$"Income 3" >= 0))
+          expect_true(all(cleaned_data_inverted$"Income 4" >= 0))
[…]

### 2024-09-27 scripts/00-simulate_data.R
fixed preamble

@@ -1,11 +1,9 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from Open Data Toronto
+# Author: Hyunje Park
+# Date: 24 September 2024
+# Contact: hyunje.park@mail.utoronto.ca 
+# Pre-requisites: Packages listed below
 
 
 #### Workspace setup ####

### 2024-09-27 scripts/01-download_data.R

@@ -3,7 +3,6 @@
 # Author: Hyunje Park
 # Date: 24 September 2024
 # Contact: hyunje.park@mail.utoronto.ca 
-# License: MIT
 # Pre-requisites: Packages listed below
 
 #### Workspace setup ####

### 2024-09-27 scripts/02-data_cleaning.R

@@ -3,7 +3,6 @@
 # Author: Hyunje Park
 # Date: 24 September 2024
 # Contact: hyunje.park@mail.utoronto.ca 
-# License: MIT
 # Pre-requisites: 01-download_data.R, packages listed below.
 
 #### Workspace setup ####

### 2024-09-27 paper/paper.qmd
fixed head() not displaying all the rows

@@
 avg_income$"Average Income in $" <- round(avg_income$"Average Income in $", 0)
-
-avg_income %>% head() %>% knitr::kable()
+avg_income %>% head(13) %>% knitr::kable()
 ```
@@
 See `Table 3` for an overview of the dataset.
+Note that all the values are rounded up and therefore may not add up to the total.
 
@@
 avg_expense$"Average Expense in $" <- abs(avg_expense$"Average Expense in $")
+
 avg_expense %>% head() %>% knitr::kable()

### 2024-09-27 scripts/00-simulate_data.R

@@ -5,13 +5,26 @@
 # Contact: hyunje.park@mail.utoronto.ca 
 # Pre-requisites: Packages listed below
 
-
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
+#### EXPECTED DATA ####
+# Columns: Income 1-11, Expense 3-6
+# 
+# All the variables contain values within the specified ranges below.
+# Determined by the average of the scenarios
+
+# number of observations
+obs = 100
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+simulated_data <-
+  tibble(
+    "Income_1" = sample(1:3000, obs, replace = TRUE)
+    
+    
+  
+  )
 
-
+print(simulated_data)
 

### 2024-09-27 scripts/02-data_cleaning.R

@@ -42,7 +42,8 @@ colnames(raw_data) = c("Category", "Item",
 
 cleaned_data <- raw_data
 
-#inverted version of raw_data, will need it for graphs in Section 3 Results
+#inverted version of raw_data, will need it for graphs in Section 3 Results,
+# as well as tests and simulations
 cleaned_data_inverted <- data.frame(t(raw_data[-1]))
 colnames(cleaned_data_inverted) <- raw_data[, 1]
 

### 2024-09-27 scripts/00-simulate_data.R
finished simulate_data.R

@@ -9,10 +9,10 @@
 library(tidyverse)
 
 #### EXPECTED DATA ####
-# Columns: Income 1-11, Expense 3-6
+# Columns: Income 1-11, Expense 1-6
 # 
 # All the variables contain values within the specified ranges below.
-# Determined by the average of the scenarios
+# Determined by the average of the scenarios.
 
 # number of observations
 obs = 100
@@ -20,11 +20,62 @@ obs = 100
 #### Simulate data ####
 simulated_data <-
   tibble(
-    "Income_1" = sample(1:3000, obs, replace = TRUE)
-    
-    
-  
+    "Income_1" = sample(1:1603, obs, replace = TRUE),
+    "Income_2" = sample(1:365, obs, replace = TRUE),
+    "Income_3" = sample(1:337, obs, replace = TRUE),
+    "Income_4" = sample(1:157, obs, replace = TRUE),
+    "Income_5" = sample(1:8, obs, replace = TRUE),
+    "Income_6" = sample(1:480, obs, replace = TRUE),
+    "Income_7" = sample(1:48, obs, replace = TRUE),
+    "Income_8" = sample(1:121, obs, replace = TRUE),
+    "Income_9" = sample(1:43, obs, replace = TRUE),
+    "Income_10" = sample(1:45, obs, replace = TRUE),
+    "Income_11" = sample(1:4, obs, replace = TRUE),
+    "Expense_1" = sample(1:24, obs, replace = TRUE),
+    "Expense_2" = sample(1:81, obs, replace = TRUE),
+    "Expense_3" = sample(1:3146, obs, replace = TRUE),
+    "Expense_4" = sample(1:705, obs, replace = TRUE),
+    "Expense_5" = sample(1:360, obs, replace = TRUE),
+    "Expense_6" = sample(1:460, obs, replace = TRUE),
   )
 
-print(simulated_data)
+# ensure all income/expense columns exist
+simulated_data
+names(simulated_data)
+
+# get the sum of income and expenses
+simulated_data$"sum_income" <- rowSums(subset(simulated_data, select = c("Income_1",
+                                                                     "Income_2",
+                                                                     "Income_3",
+                                                                     "Income_4",
+                                                                     "Income_5",
+                                                                     "Income_6",
+                                                                     "Income_7",
+                                                                     "Income_8",
+                                                                     "Income_9",
+                                                                     "Income_10",
+                                                                     "Income_11"
+)))
+
[…]

### 2024-09-27 paper/paper.qmd
last change - fixed typos

@@
 
-This report uses the "About Cost of living in Toronto for low-income households" [@citecityoftoronto] dataset, which is published publicly by the Social Development, Finance & Administration of Toronto on City of Toronto's Open Data Portal.
+This report uses the "About Cost of living in Toronto for low-income households" [@citeopendatatoronto] dataset, which is published publicly by the Social Development, Finance & Administration of Toronto on City of Toronto's Open Data Portal.
 This data contains 17 variables on income and expenses such as `employment income`, `GST/HST credits`, `pension benefits`, `rent`, `food`, `transportation costs` for 11 different scenarios of a low-income household.
 The 11 different scenarios of low-income household were manually selected by the City of Toronto to reflect different household characteristics that can increase the reliance on social assistance, such as type of income, single-parent household, etc.
-A detailed explanation of this 11 household scenarios can be found in the Appendix.
+A detailed explanation of this 11 household scenarios can be found in the @sec-appendix .
 The dataset is also within a monthly time period, reflecting monthly income and expenses.
@@
 # take the average
-avg_income$"Average Income in $" <- rowMeans(subset(avg_income, select = c("Household_1",
+avg_income$"Average Income ($)" <- rowMeans(subset(avg_income, select = c("Household_1",
                                                                            "Household_2",
@@
                                                                            )))
-avg_income <- avg_income %>% select("Item", "Average Income in $")
+avg_income <- avg_income %>% select("Item", "Average Income ($)")
 
 # rename the columns to something more presentable
-colnames(avg_income) <- c("Type of Income", "Average Income in $")
+colnames(avg_income) <- c("Type of Income", "Average Income ($)")
 
 #round to the nearest integer
-avg_income$"Average Income in $" <- round(avg_income$"Average Income in $", 0)
+avg_income$"Average Income ($)" <- round(avg_income$"Average Income ($)", 0)
 avg_income %>% head(13) %>% knitr::kable()
@@
 
-avg_expense %>% head() %>% knitr::kable()
+avg_expense %>% head(7) %>% knitr::kable()
 ```
@@
 ```
+
+\newpage
 
@@
 
+\newpage
+
 # References

### 2024-09-27 paper/paper.qmd
fixed a typo in paper.qmd

@@
 
+# code idea acquired from https://stackoverflow.com/questions/10212106/creating-grouped-bar-plot-of-multi-column-data-in-r
 data_long <- data %>%

### 2024-09-27 paper/paper.qmd
fixed cross references

@@
 The variables chosen for this analysis were variables for the 11 households, for each entry of the income and expense types.
-See Table 1 for an overview.
-It is noted that expenses are not visible in `Table 1` due to the long list of income that occupies the first rows.
+See @fig-preview for an overview.
+It is noted that expenses are not visible in @fig-preview due to the long list of income that occupies the first rows.
 
 ```{r}
-#| label: low-income-families-preview
-#| fig-cap: "Table 1: Sample of the Income/Expenses of Low Income Households in Toronto"
+#| label: fig-preview
+#| fig-cap: "Sample of the Income/Expenses of Low Income Households in Toronto"
 #| echo: false
@@
 
-`Table 2` shows that income from employment was the major income source for all the 11 households in Toronto by a large margin, at approximately $1603 per month.
+@fig-average-income shows that income from employment was the major income source for all the 11 households in Toronto by a large margin, at approximately $1603 per month.
 This number also falls shy below the "low-income" threshold set by the City of Toronto, which is approximately $2208.5 [@citecityoftoronto].
@@
 ```{r}
-#| label: low-income-families-average-income
-#| fig-cap: "Table 2: Average Income of Low Income Households in Toronto"
+#| label: fig-average-income
+#| fig-cap: "Average Income of Low Income Households in Toronto"
 #| echo: false
@@
 The huge cost rent highlights a need for further investigation and is discussed in @sec-discussion.
-See `Table 3` for an overview of the dataset.
+See @fig-average-expense for an overview of the dataset.
 Note that all the values are rounded up and therefore may not add up to the total.
@@
 ```{r}
-#| label: low-income-families-average-expense
-#| fig-cap: "Table 3: Average Expenses of Low Income Households in Toronto"
+#| label: fig-average-expense
+#| fig-cap: "Average Expenses of Low Income Households in Toronto"
 #| echo: false
@@
 ```
+
 ## Difference between Income and Expense
@@
 In addition to this, a variable for their difference was also created.
-See `Table 4` for a preview of the dataset.
+See @fig-income-expense for a preview of the dataset.
 
 ```{r}
-#| label: low-income-families4
-#| fig-cap: "Table 4: Differenece of Income and Expenses of Low Income Households in Toronto"
+#| label: fig-income-expense
+#| fig-cap: "Differenece of Income and Expenses of Low Income Households in Toronto"
 #| echo: false
@@
 
-`Figure 1` shows income and expense for each household. Overall, expenses exceed income for all households. Only one household had an income that exceed expenses.
+@fig-difference shows income and expense for each household. Overall, expenses exceed income for all households. Only one household had an income that exceed expenses.
 
 ```{r}
-#| label: low-income-families5
+#| label: fig-difference
 #| fig-cap: "Total Income and Expenses of Low Income Households in Toronto"
@@
 ```
-Furthermore, `Figure 2` gives us a better understanding of the spending behind the high expenses. Rent was by far the most expensive, comprising 66% of the expenses.
+Furthermore, @fig-breakdown gives us a better understanding of the spending behind the high expenses. Rent was by far the most expensive, comprising 66% of the expenses.
 
 ```{r}
-#| label: low-income-families6
+#| label: fig-breakdown
 #| fig-cap: "Expense Breakdown of Low Income Households in Toronto"
@@
 However, looking at the results suggest some inefficiencies in spending.
-For example, in `Figure 3`, the average rent was $3146 per month for the 11 families, a big difference from the average rent for a one-bedroom condo of $2,529 in Toronto for Q2 of 2023 [@citerentalmarket].
+For example, in @fig-average-expense, the average rent was $3146 per month for the 11 families, a big difference from the average rent for a one-bedroom condo of $2,529 in Toronto for Q2 of 2023 [@citerentalmarket].
 Expenses such as this could suggest that expenses could be correlated with poor financial decisions, however more research must be conducted to draw those conclusions.
